---
title: "US Educational Inequalities"
author: "Duncan Gates"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::html_clean: 
    self_contained: true # Other options are downcute, material, readthedown, html_clean, html_docco, lockdown, https://github.com/juba/rmdformats
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    number_sections: true
    includes:
      after_body: footer.html
---

```{css}
h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

```{r echo = F, message = F, warning = F}
library(tidyverse)
library(tidycensus)
library(leaflet)
library(sf)
library(mapview)
library(viridis)
library(hrbrthemes)
theme_set(theme_ipsum())
```

# Tidycensus Capabilities

Check what variables are available from the 5 year American Community Survey in 2019.

```{r eval = F}
all_2019_variables <- load_variables(2019, "acs5")
```

Obtain data such as median home value easily:

```{r message = F, warning = F, echo = F}
# Sys.setenv(CENSUS_API_KEY = "aa7dad57625fd3a7f42a8066e1f3a2ea3ff31ed7")
# Sys.getenv("CENSUS_API_KEY") # Checks for API key
# us <- unique(fips_codes$state)[1:51] # gets list of states by abbreviation
# 
# totalpop_sf <- map_df(us, function(x) {
#   get_acs(geography = "tract", variables = "B01003_001",
#           state = x, geometry = TRUE, shift_geo = T) # Do without geometry = T for more just population information
# })

totalpop_sf <- get_acs(geography = "state", variables = "B01003_001", geometry = TRUE, shift_geo = T)

# str(totalpop_sf)
# pal <- colorQuantile("Blues", domain = oregon_pop$estimate, n = 10) # blue color palette for percentiles
pal <- colorNumeric(palette = "Blues", domain = totalpop_sf$estimate) # Numeric color palette

totalpop_sf %>% # to filter by state do filter(grepl("Oregon", NAME)) %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet(width = "100%") %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(popup = ~ str_extract(NAME, "^([^,]*)"),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal(estimate)) %>%
  addLegend("bottomright", 
            pal = pal, 
            values = ~ estimate,
            title = "Median Home Value",
            labFormat = labelFormat(prefix = "$"),
            opacity = 1)
```


## County level data

```{r results='hide', echo = F, message = F, warning = F}
racevars <- c(White = "B02001_002", 
              Black = "B02001_003", 
              Asian = "B02001_005",
              Hispanic = "B03003_003")
harris <- get_acs(geography = "tract",
                  variables = racevars, 
                  year = 2019,
                  state = "TX",
                  county = "Harris County",
                  geometry = TRUE,
                  summary_var = "B02001_001") # JUST MIX WITH EDUCATION
```

```{r message = F, warning = F, echo = F, error = F}
harris %>%
  mutate(Percent = 100 * (estimate / summary_est)) %>%
  ggplot(aes(fill = Percent, color = Percent)) +
  facet_wrap(~ variable) +
  geom_sf() +
  scale_fill_viridis_c(direction = -1) +
  scale_color_viridis_c(direction = -1) +
  labs(title = "Racial geography of Harris County, Texas",
       caption = "Source: American Community Survey") +
  theme_void()
```

## Using mapview

```{r, results = 'hide', message = F, warning = F, echo = F, error = F}
ny <- get_acs(geography = "tract",
              variables = "B19013_001",
              state = "NY",
              county = "New York",
              geometry = TRUE)
```

```{r message = F, warning = F, echo = F, error = F}
mapview(ny, zcol = "estimate", legend = TRUE)
```



# Getting Educational Data

## The Actual Educational Data

Digging into 2016-2019 census data by using each one, this takes significant time on the census website learning the codes and API formatting.

```{r}
get_historic_acs <- function (variables, 
                              geography, 
                              year, 
                              summary_var = NULL) {
y <- list()
for (i in 1:length(year)) {
  y[[i]] <- lapply(geography, function (x) {
      tidycensus::get_acs (geography = x, 
                           variables = variables, 
                           summary_var = summary_var, 
                           output = "tidy", 
                           year = year[i])}) %>%
      bind_rows() %>% 
      mutate(year = year[i]) } 
  y %>% bind_rows() } # Function to get all data instead of just most recent

year_range <- 2016:2019

geos_inc <- c("us", 
              "metropolitan statistical area/micropolitan statistical area", 
              "state")

variable <- c('DP02_0059P', 
              'DP02_0060P', 
              'DP02_0061P', 
              'DP02_0062P', 
              'DP02_0063P', 
              'DP02_0064P', 
              'DP02_0065P')

ed_labels <- c('Less than 9th Grade', 
               '9th to 12th grade, no diploma', 
               'High school graduate', 
               'Some college, no degree', 
               "Associate's degree", 
               "Bachelor's degree", 
               'Grad/pro degree')

ed_level <- c(1:7)

ed_table <- as_tibble(bind_cols(variable, ed_level, ed_labels), stringsAsFactors = FALSE)

ed_data <- get_historic_acs(variables = variable, 
                            geography = geos_inc, 
                            year = year_range)
ed_data <- ed_data %>% 
  mutate(variable = case_when(variable == "DP02_0059P" ~ 'Less than 9th Grade',
                              variable == "DP02_0060P" ~ '9th to 12th grade, no diploma',
                              variable == "DP02_0061P" ~ 'High school graduate',
                              variable == "DP02_0062P" ~ 'Some college, no degree',
                              variable == "DP02_0063P" ~ 'Associate\'s degree',
                              variable == "DP02_0064P" ~ 'Bachelor\'s degree',
                              variable == "DP02_0065P" ~ 'Grad/pro degree'))
```


## Secondary Data for Plotting and Comparison

```{r}
variable <-c("DP02_0011P", 
             "DP02_0067P",
             "DP02_0069P",
             "DP02_0090P", 
             "DP02_0092P", 
             "DP02_0111P", 
             "DP03_0005P", 
             "DP03_0021P", 
             "DP03_0028P",
             "DP03_0088",
             "DP03_0096P",
             "DP03_0128P")

label <- c("%Householders living alone", 
           "%Bachelor's degree or higher", 
           "%Civilian veterans",
           "%Born different state", 
           "%Foreign born",
           "%Speak English only @ home", 
           "%Civilian LF - Unemployed",  
           "%Public trans to work", 
           "%Service occupations", 
           "$Per capita income", 
           "%Health insurance", 
           "%Below FPL - All people")
dp_table <- as.data.frame(cbind(variable, label))

dp_data <- get_historic_acs(variables=variable, 
                            geography = geos_inc, 
                            year = year_range) # REMINDER TO GGANIMATE LAST 5 YEARS

race_variables <- c("B02001_001",
                    "B02001_002",
                    "B02001_003",
                    "B02001_004",
                    "B02001_005")

race_labels <- c("Total",
                 "White",
                 "Black",
                 "American Indian",
                 "Asian")

race_table <- as_tibble(bind_cols(race_variables, race_labels))

race_data <- get_historic_acs(variables = race_variables, 
                            geography = geos_inc, 
                            year = year_range)

race_data <- race_data %>% mutate()

# Easier way to do it
race_data2 <- get_acs(geography = "county",
                  variables = racevars,
                  year = 2019,
                  geometry = TRUE,
                  summary_var = "B02001_001")




```










