---
title: "Sentiment Analysis of Kanye's Lyricism"
date: "[Return to Site](https://duncangates.me)"
output:
  rmdformats::html_docco:
    self_contained: true # Other options are downcute, material, readthedown, html_clean, html_docco, lockdown, https://github.com/juba/rmdformats
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: kate
    number_sections: false
    includes: 
      after_body: footer.html
---

<style type="text/css">
date {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(spotifyr)
library(plyr)
library(tidyverse)
library(httr)
library(rvest)
library(stringr)
library(ggthemes)
library(tidytext)
library(wordcloud)
library(ggridges)
library(wesanderson)
library(yarrr)
library(knitr)
library(kableExtra)
library(DT)
library(radarchart)
library(lubridate)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=100)
htmltools::tagList(rmarkdown::html_dependency_font_awesome()) # Needed so fa's in footer will show
# set up Spotify client ID and client secret
Sys.setenv(SPOTIFY_CLIENT_ID = '383cfac3d8434244a38c4e279a04ce47')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '486c6f539f8e45a2b82f03a30cfed26f')
token <- "QZWrzUmE0Kg2QvTjynbk2gqnIV9S_FaE_Cax36z54N_FcvDgZHGHB2sngC5-JYiD"

# Super sick site here https://www.color-hex.com/color-palette/90639
# Can also get them from adobe color
colors_cd <- c("#411218", "#e8982e", "#8087b5", "#ddb890")
colors_lr <- c("#633119", "#7398aa", "#c3ab14", "#373123", "#b18662")
colors_grad <- c("#221e1f", "#ffffff", "#b9b353", "#853082", "#b25186")
colors_808s <- c("#de6656", "#6e4b34", "#a89c9a", "#b72124", "#c2cece")
colors_mbdtf <- c("#EC2440", "#BD2517", "#454F58", "#C9A48B", "#845242")
colors_yeez <- c("#D8D9DE", "#FDFEFF", "#D9F7F7", "#F9FCDF", "#F20505")
colors_tlop <- c("#7acec2", "#e6dfcc", "#412517", "#a75f3b", "#f58b57")
colors_ye <- c("#172940", "#7A8CA4", "#08F42A", "#11DA2D", "#F2EEE5")
kanyealbums <- c("ye", "The Life Of Pablo", "Yeezus", "Watch The Throne", "My Beautiful Dark Twisted Fantasy", "808s & Heartbreak", "Graduation", "Late Registration", "The College Dropout"
)
theme_set(theme_minimal())
```


```{r echo = F}
# options("digits" = 3)
# readr::write_rds(spotify_genius, "/Users/dunk/Github/SpotifyStats/KanyeFinalAnalysis/main_data.csv")
spotify_genius <- readr::read_rds("/Users/dunk/Projects/NewWebsite.io/rmd/Kanye/main_data.csv")
```

# Plotting the Music

First, let's take a look at valence, which by Spotify's definition is an attribute "that describes the musical positiveness conveyed by a track. Tracks with high valence sound more positive (happy, cheerful, euphoric), while tracks with low valence sound more negative (sad, depressed, angry)."

<center>
```{r fig.width = 14}
spotify_genius %>% ggplot(aes(x = valence, y = ordered_albums, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 0.9) +
  scale_fill_gradient(low = "white", high = "deeppink") +
  theme_fivethirtyeight() +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  xlim(0,1) +
  ggtitle("Kanye's Valence By Album: A Clear Path of Misery") +
  theme(legend.position = "none")
```

Before beginning the sentiment analysis it seems useful to look at what the songs mean using Spotify's metrics to evaluate the emotions of a song, the following plots investigate how valence (x-axis) interacts with energy (y-axis), essentially producing a 4 quadrant graph where from quadrant 1-4 we have Anger, Happiness, Sadness, and Calmness. The colors on the labels in these graphs match the colors of the album covers just for fun, they don't actually mean anything.

<center>
```{r, fig.width = 14}
colors_cd <- c("#411218", "#e8982e", "#8087b5", "#ddb890",
               "#411218", "#e8982e", "#8087b5", "#ddb890",
               "#411218", "#e8982e", "#8087b5", "#ddb890",
               "#411218", "#e8982e", "#8087b5", "#ddb890",
               "#411218", "#e8982e", "#8087b5", "#ddb890",
               "#411218")
spotify_genius %>% 
  dplyr::filter(album_name == "The College Dropout") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_cd) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme_fivethirtyeight() +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("The College Dropout: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

Filler Text 2.

<center>
```{r, fig.width = 14}
colors_lr <- c("#633119", "#7398aa", "#c3ab14", "#373123", "#b18662",
               "#633119", "#7398aa", "#c3ab14", "#373123", "#b18662",
               "#633119", "#7398aa", "#c3ab14", "#373123", "#b18662",
               "#633119", "#7398aa", "#c3ab14", "#373123", "#b18662",
               "#633119", "#7398aa")
spotify_genius %>% 
  dplyr::filter(album_name == "Late Registration") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_lr) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  theme_fivethirtyeight() +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("Late Registration: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

Filler Text 3.

<center>
```{r, fig.width = 14}
colors_grad <- c("#221e1f", "#ffffff", "#b9b353", "#853082", "#b25186",
                 "#221e1f", "#ffffff", "#b9b353", "#853082", "#b25186",
                 "#221e1f", "#ffffff", "#b9b353", "#853082")
spotify_genius %>% 
  dplyr::filter(album_name == "Graduation") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_grad) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  theme_fivethirtyeight() +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("Graduation: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

Filler Text 4.

<center>
```{r, fig.width = 14}
colors_808s <- c("#de6656", "#6e4b34", "#a89c9a", "#b72124", "#c2cece",
                 "#de6656", "#6e4b34", "#a89c9a", "#b72124", "#c2cece",
                 "#de6656", "#6e4b34")
spotify_genius %>% 
  dplyr::filter(album_name == "808s & Heartbreak") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_808s) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  theme_fivethirtyeight() +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("808s & Heartbreak: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

Filler Text 5.

<center>
```{r, fig.width = 14}
colors_mbdtf <- c("#EC2440", "#BD2517", "#454F58", "#C9A48B", "#845242",
                  "#EC2440", "#BD2517", "#454F58", "#C9A48B", "#845242",
                  "#EC2440", "#EC2440")
spotify_genius %>% 
  dplyr::filter(album_name == "My Beautiful Dark Twisted Fantasy") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_mbdtf) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  theme_fivethirtyeight() +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("My Beautiful Dark Twisted Fantasy: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```


Filler Text 6.

<center>
```{r, fig.width = 14}
colors_yeez <- c("#D8D9DE", "#FDFEFF", "#D9F7F7", "#F9FCDF", "#F20505",
                 "#D8D9DE", "#FDFEFF", "#D9F7F7", "#F9FCDF", "#F20505")
spotify_genius %>% 
  dplyr::filter(album_name == "Yeezus") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_yeez) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  theme_fivethirtyeight() +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("Yeezus: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

Filler Text 7.

<center>
```{r, fig.width = 14}
colors_tlop <- c("#7acec2", "#e6dfcc", "#412517", "#f58b57", "#f58b57",
                 "#7acec2", "#f58b57", "#412517", "#a75f3b", "#f58b57",
                 "#f58b57", "#e6dfcc", "#412517", "#a75f3b", "#f58b57",
                 "#412517", "#e6dfcc", "#412517", "#a75f3b", "#f58b57")
spotify_genius %>% 
  dplyr::filter(album_name == "The Life Of Pablo") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_tlop) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  theme_fivethirtyeight() +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("The Life Of Pablo: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

Filler Text 8.

<center>
```{r, fig.width = 14}
colors_ye <- c("#172940", "#7A8CA4", "#172940", "#11DA2D", "#F2EEE5",
               "#7A8CA4", "#7A8CA4")
spotify_genius %>% 
  dplyr::filter(album_name == "ye") %>%
  ggplot(aes(x = valence, y = energy)) +
  ggrepel::geom_label_repel(aes(label = track_name), fill = colors_ye) +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  annotate(geom = "text", label = "Happy", x = 0.9, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Angry", x = 0.1, y = 0.9, fontface = 2) +
  annotate(geom = "text", label = "Calm", x = 0.9, y = 0.1, fontface = 2) +
  annotate(geom = "text", label = "Sad", x = 0.1, y = 0.1, fontface = 2) +
  scale_fill_gradient(low = "white", high = "maroon3") +
  theme_fivethirtyeight() +
  labs(x = "Valence (Happiness)", y = "Energy", color = "Album", alpha = "Tempo") + scale_alpha(guide = "none") +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  ggtitle("ye: Kanye's Energy and Valence") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
```

Graphing is great but if you want to look at each song in more detail here is a table color coded by album.


```{r}
# table: album by mean valence
spotify_genius %>%
  dplyr::group_by(factor(album_name, levels = c(kanyealbums))) %>%
  dplyr::summarise(round(mean(valence), 4)) %>%
  arrange(desc(`factor(album_name, levels = c(kanyealbums))`)) %>%
  dplyr::rename(Album = `factor(album_name, levels = c(kanyealbums))`, `Mean Valence` = `round(mean(valence), 4)`) %>%
  datatable() %>%
  formatStyle(target = "row",
              columns = c("Album"),
              backgroundColor = styleEqual(c("The College Dropout", "Late Registration",
                                             "Graduation", "808s & Heartbreak",
                                             "My Beautiful Dark Twisted Fantasy", "Yeezus",
                                             "The Life Of Pablo", "ye"),
                c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4")))
```

Let's check out all the songs organized by valence, once again color coded by album. The Life of Pablo, College Dropout, and Graduation all feature highest here.

```{r}
# Songs organized by valence
spotify_genius %>%
  select(track_name, album_name, valence) %>%
  arrange(-valence) %>%
  dplyr::rename(Album = album_name,
         Track = track_name,
         Valence = valence) %>%
  dplyr::relocate(Album, .before = Track) %>%
  datatable() %>%
  formatStyle(target = "row",
              columns = c("Album"),
              backgroundColor = styleEqual(c("The College Dropout", "Late Registration",
                                             "Graduation", "808s & Heartbreak",
                                             "My Beautiful Dark Twisted Fantasy", "Yeezus",
                                             "The Life Of Pablo", "ye"),
                c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4")))
```

Now a plot of valence, danceability, and energy as functions of album year, which for Kanye is just his album discography.

<center>
```{r}
pirateplot(valence + danceability + energy ~ album_release_year, spotify_genius,
           pal = c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4"),
           xlab = "Album", ylab = "Sonic Score",
           theme = 1, point.o = 0.7, avg.line.o = 1, jitter.val = .05,
           bty = "n", cex.axis = 0.6, xaxt = "n",
           bean.f.o = 0.5)
axis(1, cex.axis = 0.6, lwd = 0, labels = F)
legend("top", c("College Dropout", "Late Registration",
                     "Graduation", "808s and Heartbreaks"), bty = "n", cex = 0.6, fill = c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4"))
legend("topright", c("My Beautiful Dark Twisted Fantasy",
                     "Yeezus", "The Life Of Pablo", "ye"), bty = "n", cex = 0.6, fill = c("#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4"))
```
</center>

Sonic Score Table.

```{r}
spotify_genius %>%
  dplyr::mutate(sonic_score = valence + danceability + energy) %>%
  select(album_name, track_name, sonic_score) %>%
  arrange(desc(sonic_score)) %>%
  dplyr::group_by(album_name) %>%
  dplyr::rename(Album = album_name, `Sonic Score` = sonic_score, Track = track_name) %>%
  datatable() %>%
  formatStyle(target = "row",
    columns = c("Album"),
              backgroundColor = styleEqual(c("The College Dropout", "Late Registration",
                                             "Graduation", "808s & Heartbreak",
                                             "My Beautiful Dark Twisted Fantasy", "Yeezus",
                                             "The Life Of Pablo", "ye"),
                c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4")))
```

This table has the mean of all of Spotify's metrics by album, which honestly seems to show a fairly consistent pattern for Kanye's discography.


```{r}
spotify_genius %>%
  dplyr::group_by(factor(album_name, levels = c(kanyealbums))) %>%
  dplyr::summarise(Danceability = mean(danceability),
                   Energy = mean(energy),
                   Tempo = mean(tempo),
                   Liveness = mean(liveness),
                   Instrumentalness = mean(instrumentalness),
                   Loudness = mean(loudness),
                   Acousticness = mean(acousticness),
                   Valence = mean(valence),
                   Duration = mean(duration_ms),
                   Tracks = n()) %>%
  arrange(desc(`factor(album_name, levels = c(kanyealbums))`)) %>%
  dplyr::rename(Album = `factor(album_name, levels = c(kanyealbums))`) %>%
  datatable(options = list(scrollX = TRUE)) %>%
  formatStyle(target = "row",
              columns = c("Album"),
              backgroundColor = styleEqual(c("The College Dropout", "Late Registration",
                                             "Graduation", "808s & Heartbreak",
                                             "My Beautiful Dark Twisted Fantasy", "Yeezus",
                                             "The Life Of Pablo", "ye"),
                c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4")))

```


# Sentiment Analysis

```{r echo = F}
# tokenized and cleaned datasets of lyrics for textual analysis
tidy_kanye <- spotify_genius %>% unnest_tokens(word, lyrics)
tidier_kanye <- tidy_kanye %>% anti_join(rbind(stop_words[1], "uh", "yeah", "hey", "baby", "ooh", "wanna", "gonna", "ah", "ahh", "ha", "la", "mmm", "whoa", "haa"))
tidier_kanye$word[tidier_kanye$word == "don" | tidier_kanye$word == "didn"] <- NA
tidier_kanye$word[tidier_kanye$word == "ain"] <- NA
tidier_kanye$word[tidier_kanye$word == "isn"] <- NA
tidier_kanye$word[tidier_kanye$word == "usin"] <- "using"
tidier_kanye$word[tidier_kanye$word == "wouldn"] <- "wouldn't"
tidier_kanye$word[tidier_kanye$word == "couldn"] <- "couldn't"
tidier_kanye$word[tidier_kanye$word == "shouldn"] <- "shouldn't"
tidier_kanye$word[tidier_kanye$word == "won"] <- "won't"
tidier_kanye$word[tidier_kanye$word == "ve" | tidier_kanye$word == "ll"] <- NA
tidier_kanye$word[tidier_kanye$word == "ileft"] <- "left"
```

Wordcloud of the most frequent words in Kanye's songs after filtering out common meaningless words like the, a, etc.

<center>
```{r}
word_count <- tidier_kanye %>%
  dplyr::count(word, sort = TRUE) %>%
  dplyr::mutate(word = reorder(word, n)) %>%
  dplyr::ungroup()

wordcloud(words = word_count$word, freq = word_count$n,
          max.words=200, random.order=FALSE,
          colors= c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4"))
```
</center>

How many tracks does the word "kanye" appear in, and how often? Obviously a lot in "I Love Kanye", but surprisingly enough only once in "Wake Up Mr. West", and not even that much overall, I would not be surprised if most artists had similar amounts of self-references throughout their discography. Since he often refers to himself as ye, or yeezus, I also include those in the table count, but neither of those even show up that much, yeezus only appearing once in "Highlights" and "I am a God" each.


```{r}
# how many tracks does the word "kanye" appear in?
kanye_freq <- tidier_kanye %>%
  select(track_name, word) %>%
  dplyr::filter(word == "kanye" | word == "ye" | word == "yeezus") %>%
  dplyr::group_by(track_name) %>%
  dplyr::summarise(plyr::count(word)) %>%
  dplyr::rename(Word = x, Count = freq, Track = track_name) %>%
  dplyr::arrange(desc(Count))
datatable(kanye_freq) %>%
  formatStyle(
  'Word',
  target = 'row',
  backgroundColor = styleEqual(c('kanye'), c('lightblue'))
)
  # Find an all data table where this can be used
  # formatStyle(names(kanye_freq),
  # background = styleColorBar(range(kanye_freq), 'lightblue'),
  # backgroundSize = '98% 88%',
  # backgroundRepeat = 'no-repeat',
  # backgroundPosition = 'center')
```

Wordcloud for just the album ye, you can see why sentiment analysis would pick this up as deeply depressed/angry/sad.

<center>
```{r}
# wordcloud: make this function work by album later
word_count_ts <- tidier_kanye %>%
  dplyr::filter(album_name == "ye") %>%
  dplyr::count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  ungroup()

wordcloud(words = word_count_ts$word, freq = word_count_ts$n,
          max.words=25, random.order=FALSE,
          colors= colors_ye)
```
</center>

Wordcloud for 808s & Heartbreak, the issue is clear once again, although this time it could literally just because Kanye says "Amazing" A LOT in the song amazing, in fact he says it 55 times, which is the third most frequent by song out of all of his discography. Coming just after "bam" which isn't sang by him in Famous, and "ey" also in Famous, it can really be argued that it is the largest personal use in his discography.

<center>
```{r}
# library(gt)
# tidier_kanye %>%
#   dplyr::group_by(track_name) %>%
#   dplyr::count(word, sort = T) %>%
#   gt() %>%
#   tab_header(
#     title = "Kanye Word Counts by Track"
#   ) # FIGURE OUT WHY THIS DOESN'T WORKS
word_count_ts_2 <- tidier_kanye %>%
  dplyr::filter(album_name == "808s & Heartbreak") %>%
  dplyr::count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  ungroup()

wordcloud(words = word_count_ts_2$word, freq = word_count_ts_2$n,
          max.words=25, random.order=FALSE,
          colors= colors_808s)
```
</center>



Wordcloud for My Beautiful Dark Twisted Fantasy, not much to interpret here there is a lot going on, but if you've heard this album this makes sense.

<center>
```{r}
word_count_rep <- tidier_kanye %>%
  dplyr::filter(album_name == "My Beautiful Dark Twisted Fantasy") %>%
  dplyr::count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  ungroup()

wordcloud(words = word_count_rep$word, freq = word_count_rep$n,
          max.words=25, random.order=FALSE,
          colors= colors_mbdtf)
```
</center>

Lexical diversity is fairly consistently in the middle ranges, although there are some outliers in The Life of Pablo and The College Dropout that should be investigated.

<center>
```{r}
# more cleaning, can be done earlier
tidier_kanye$album_release_year <- as.character(tidier_kanye$album_release_year)
tidier_kanye$album_release_year <- as.numeric(substr(tidier_kanye$album_release_year, 1, 4))

tidy_kanye$album_release_year <- as.character(tidy_kanye$album_release_year)
tidy_kanye$album_release_year <- as.numeric(substr(tidy_kanye$album_release_year, 1, 4))


# creating a "lexical diversity" dataset
lexical_diversity <- tidy_kanye %>% 
  dplyr::group_by(track_name, album_release_year) %>%
  dplyr::mutate(lex_div = length(unique(word))/length(word)) %>%
  select(track_name, lex_div, album_release_year) %>%
  distinct()


# lexical diversity plot
pirateplot(lex_div ~ album_release_year, lexical_diversity,
           pal = c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4"),
           xlab = "Album", ylab = "Lexical Diversity",
           theme = 0, point.o = 0.5, avg.line.o = 1, jitter.val = .05,
           bty = "n", cex.axis = 0.6, xaxt = "n")
axis(1, cex.axis = 0.6, lwd = 0, labels = F)
legend("top", c("College Dropout", "Late Registration",
                     "Graduation", "808s and Heartbreaks"), bty = "n", cex = 0.6, fill = c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4"))
legend("topright", c("My Beautiful Dark Twisted Fantasy",
                     "Yeezus", "The Life Of Pablo", "ye"), bty = "n", cex = 0.6, fill = c("#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4"))
```
</center>


```{r}
# least lexically diverse tracks
tidy_kanye %>% dplyr::group_by(track_name, album_name) %>%
  dplyr::mutate(lex_div = length(unique(word))/length(word)) %>%
  select(track_name, lex_div, album_name) %>%
  arrange(lex_div) %>%
  distinct() %>%
  dplyr::rename(Album = album_name) %>%
  datatable() %>%
  formatStyle(target = "row",
              columns = c("Album"),
              backgroundColor = styleEqual(c("The College Dropout", "Late Registration",
                                             "Graduation", "808s & Heartbreak",
                                             "My Beautiful Dark Twisted Fantasy", "Yeezus",
                                             "The Life Of Pablo", "ye"),
                c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4")))
```

Now for the real fun, sentiment analysis using NRC, AFINN, and BING, which is surprisingly straightforward.


```{r echo = F}
# joining the tokenized, tidied lyric dataset with sentiment lexicons
kanye_nrc_sub <- tidier_kanye %>%
  inner_join(get_sentiments("nrc")) %>%
  dplyr::filter(!sentiment %in% c("positive", "negative"))

kanye_AFINN <- tidier_kanye %>%
  inner_join(get_sentiments("afinn"))

kanye_bing <- tidier_kanye %>%
  inner_join(get_sentiments("bing"))
```


Sentiment scores with AFINN, looks like its not a great tool in general given that it got the exact opposite of what it should have - that 808s was his most depressed and lowest point, it would be useful in the future to try to weight instrumental components in this analysis, to try to give context to the data.

<center>
```{r}
# sentiment scores using AFINN
dim <- kanye_AFINN %>%
  dplyr::count(album_name)
kanye_AFINN %>%
  dplyr::group_by(ordered_albums) %>%
  dplyr::summarise(sum(value)) %>%
  dplyr::mutate(scaled = `sum(value)` * 229 / dim$n) %>%
  ggplot(aes(x = ordered_albums, y = scaled)) +
  geom_bar(stat = "identity", fill = c("#411218",
                         "#633119",
                         "#853082",
                         "#c2cece",
                         "#EC2440",
                         "#D8D9DE",
                         "#f58b57",
                         "#7A8CA4")) +
  coord_flip() +
  labs(title = "AFINN Scores by Album") +
  theme_fivethirtyeight() +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white")) +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5))
```
</center>

Pyramid Plot of The Life of Pablo

<center>
```{r}
kanye_pyramid <- kanye_bing %>%
  unique() %>%
  dplyr::group_by(track_name, sentiment, album_name) %>%
  dplyr::count(track_name, sentiment)

for(i in 1:24) {
  if(kanye_pyramid$sentiment[i] == "negative")
    kanye_pyramid$n[i] <- -kanye_pyramid$n[i]
}

kanye_pyramid %>%
  dplyr::mutate(sentiment = str_to_title(sentiment)) %>%
  dplyr::filter(album_name == "The Life Of Pablo") %>%
  ggplot(aes(x = track_name, y = n, fill = sentiment)) +
  geom_bar(subset = .(sentiment == "positive"), stat = "identity") +
  geom_bar(subset = .(sentiment == "negative"), stat = "identity") +
  scale_y_continuous() +
  coord_flip() +
  ggtitle("The Life Of Pablo Bing Scores") +
  labs(fill = "Sentiment") +
  theme_fivethirtyeight() +
  scale_fill_manual(values = c("#f58b57", "#412517")) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(plot.background = element_rect(fill = "white"))
```
</center>

Radar chart.

<center>
```{r}
sentiment_nrc <- kanye_nrc_sub %>%
  dplyr::group_by(ordered_albums, sentiment) %>%
  dplyr::count(ordered_albums, sentiment) %>%
  select(ordered_albums, sentiment, sentiment_total = n) %>%
  mutate(sentiment = str_to_title(sentiment))

album_nrc <- kanye_nrc_sub %>%
  dplyr::count(ordered_albums) %>%
  select(ordered_albums, album_total = n)

radar_chart <- sentiment_nrc %>%
  inner_join(album_nrc, by = "ordered_albums") %>%
  mutate(percent = round((sentiment_total/album_total * 100), 3)) %>%
  select(-sentiment_total, -album_total) %>%
  spread(ordered_albums, percent)

radar_chart <- radar_chart[c(2,7,5,8,4,3,1,6), c(1, 7:2)]

# Cool Radar Chart of Spotify Data
chartJSRadar(radar_chart, polyAlpha = 0.2, lineAlpha = 0.8, maxScale = 25,
             colMatrix = matrix(c(0, 255, 255, 255, 185, 15, 139, 0, 139,
                                  255, 0, 0, 201, 167, 198, 0, 0, 0), byrow = F, nrow = 3))
```
</center>

Most common words in song.

<center>
```{r}
text_wordcounts <- tidy_kanye %>% select(track_name, album_name, word)
text_wordcounts$word <- as.factor(text_wordcounts$word)
text_wordcounts <- text_wordcounts %>% dplyr::group_by(word) %>% dplyr::summarise(n = n()) %>% ungroup

# Ranking most common words in Kanye Songs
text_wordcounts %>% dplyr::mutate(word = reorder(word, n)) %>% anti_join(stop_words) %>%
  dplyr::filter(n > 100) %>%
  ggplot(aes(word, n, fill = -n)) + 
  geom_col() + 
  coord_flip() +
  scale_fill_distiller(palette = "Blues") +
  labs(x = "Word", y = "Count", title = "Most Frequent Words in Kanye's Songs") +
  geom_text(aes(label = n), hjust = 1.2, color = "white", fontface = "bold") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(face = "bold", color = "skyblue", size = 12),
        axis.title.y = element_blank(),
        legend.position = "none",
        axis.text.y = element_text(face = "bold", color = "skyblue", size = 10))
```
</center>


