
## The NYC Taxi Data

Almost everyone has heard of the NYC taxi data at this point, in its current form it is about 24 columns and just under 1.7 billion rows. Each row represents a ride that occurred somewhere between 2009 and 2022. The important columns in this visualization are

- `pickup_longitude` (double): Longitude data for the pickup location
- `pickup_latitude` (double): Latitude data for the pickup location
- `dropoff_longitude` (double): Longitude data for the dropoff location
- `dropoff_latitude` (double): Latitude data for the dropoff location

## Libraries

There are the libraries used for tranforming and visualizing.

```{r}
library(arrow)
library(dplyr)
library(tictoc)
library(tidyr)
library(ggplot2)
```

## Loading the Data 

This transfers the nearly 70 gigabytes of taxi data to my computer.

```{r eval = F}
copy_files(
  from = s3_bucket("ursa-labs-taxi-data-v2"),
  to = "~/Downloads/nyc-taxi"
)
```

The datasets can subsequently be opened as such.

```{r}
nyc_taxi_tiny <- open_dataset("~/Downloads/nyc-taxi-tiny/")
nyc_taxi <- open_dataset("~/Downloads/nyc-taxi/")
```

## Checking it out

```{r}
glimpse(nyc_taxi_tiny)
```

## Plotting a million rows

```{r}
tic()
nyc_pickups <- nyc_taxi_tiny |>
  select(pickup_longitude, pickup_latitude) |>
  filter(
    !is.na(pickup_longitude),
    !is.na(pickup_latitude)
  ) |>
  collect()
toc()
```


Let's check out `nyc_pickups`

```{r}
glimpse(nyc_pickups)
```

This goes pretty quick now

```{r}
x0 <- -74.05 # minimum longitude to plot
y0 <- 40.6   # minimum latitude to plot
span <- 0.3  # size of the lat/long window to plot

tic()
pic <- ggplot(nyc_pickups) +
  geom_point(
    aes(pickup_longitude, pickup_latitude), 
    size = .2, 
    stroke = 0, 
    colour = "#800020"
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_void() +
  coord_equal(
    xlim = x0 + c(0, span), 
    ylim = y0 + c(0, span)
  )
pic
toc()
```


## Plotting a billion rows

```{r}
tic()
pixels <- 4000
pickup <- nyc_taxi |>
  filter(
    !is.na(pickup_longitude),
    !is.na(pickup_latitude),
    pickup_longitude > x0,
    pickup_longitude < x0 + span,
    pickup_latitude > y0,
    pickup_latitude < y0 + span
  ) |>
  mutate(
    unit_scaled_x = (pickup_longitude - x0) / span,
    unit_scaled_y = (pickup_latitude - y0) / span,
    x = as.integer(round(pixels * unit_scaled_x)), 
    y = as.integer(round(pixels * unit_scaled_y))
  ) |>
  count(x, y, name = "pickup") |>
  collect()
toc()
```

My laptop solves this in 27.51 seconds, again lets take a look at the resulting dataframe.

```{r}
glimpse(pickup)
```


```{r}
tic()
grid <- expand_grid(x = 1:pixels, y = 1:pixels) |>
  left_join(pickup, by = c("x", "y")) |>
  mutate(pickup = replace_na(pickup,  0))
toc()
```

```{r}
tic()
pickup_grid <- matrix(
  data = grid$pickup,
  nrow = pixels,
  ncol = pixels
)
toc()
```

```{r}
render_image <- function(mat, cols = c("white", "#800020")) {
  op <- par(mar = c(0, 0, 0, 0))
  shades <- colorRampPalette(cols)
  image(
    z = log10(t(mat + 1)),
    axes = FALSE,
    asp = 1,
    col = shades(256),
    useRaster = TRUE
  )
  par(op)
}
```

## Rendering and saving

```{r}
tic()
png(file = here::here("imgs/taxi_viz_billion_rows.png"),
    bg = "#002222",
    width = 4000,
    height = 4000)
render_image(pickup_grid, cols = c("#002222", "white", "#800020"))
dev.off()
toc()
```

